### 2022-04-20 outputs/paper/paper.Rmd
initial setup

---
title: "Disagreement Regarding 2016 Presidential Election in the United States^[Code and data are available at: https://github.com/SANGWOONG-LEE/2016electionanalysis]"
subtitle: "Possible factors effecting 2016 election"
author: 
  - Young Suk
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "This paper has been examined possible variables that influenced U.S citizens' choice for the 2016 U.S presidential election. We have found that there exists significantly different opinions regarding gender, wealth, race and age. This is important in a sesne that it shows thought trends as well as crucial decision indicators for voting in U.S.,With further survey at the end of this paper, we are going to explore whether this is happening only in U.S or not."
output:
  bookdown::pdf_document2:
    fig_caption: yes
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(dplyr)
library(knitr)
```

# Introduction
Environmental issues around the world as been very important part of modern society as many changes in the environment such as climate change, water and air pollution started to be very noticeable. While the concerns regarding the state of the environment has become popular issue, it is still is not

Donald Trump won the 2016 United States presidential election. From what we remembered, it was quite surprising considering his image outside of U.S or even in U.S. Also, before the election, Hillary Clinton was ahead in most surveys as well. There is a famous quote that we learn from the past. We thought that even though we are living in 2022, it would be meaningful to analyze the election so that we get to know what U.S citizens thought, or in broader sense, thoughts trends in nowadays and what factors we considered most important to decide our representative.

This paper is mainly going to explore possible variables that potentially had impacts on decision regarding the election as well as their analysis. However, making a choice for  which variables affected U.S citizens to lean on Trump side or Clinton side was not easy because there is nothing in this world absolute. Thus, as we mentioned above, we are going to focus on general flows of results instead of defining variables as absolute factors of the demarcation. Following that, this paper has structures: Data, Results, Discussion, Appendix and supplementary survey.

The dataset was obtained from U.S GSS, and only picked and used the variables that are very related to the subject of this study. Throughout the analysis of relation between the variables and respondents' votes on the election, we could observe that people who had conservative thoughts on equality and welfare preferred Trump, and people who did not preferred Clinton. This may sound obvious considering their parties. In the result session, reader can observe that what specific areas or fields the respondents considered important. We believe that this situation is not only happening in U.S, but it is also happening all over the world. This claim is going to be conducted with supplementary survey session.

This report was created using `R` [@citeR] and its libraries `tidyverse` [@tidyverse], `dplyr` [@dplyr], were used to process and clean the data, and `ggplot` [@ggplot] `bookdown` [@bookdown] and `knitr` [@knitr] was used to format this report.

# Data
```{r, include=FALSE, echo=FALSE}
# Read the reduced and cleaned data, modification done in data-preparation.r file
reduced_data <- read_csv("outputs/data/prepared_gss.csv")
reduced_data$pres16 <- factor(reduced_data$pres16, levels = c("Clinton","Trump","Other","Did not vote"))
# Change the order of variables for graphing purposes
reduced_data$Family_income <- factor(reduced_data$Family_income, 
                                     levels = c("Under $10,000","$10,000 ~ $19,999","$20,000 ~ $39,999","$40,000 ~ $59,999", 
                                                "$60,000 ~ $89,999","$90,000 ~ $129,999","$130,000 ~ $169,999","$170,000 or over"))
reduced_data$Immigration <- factor(reduced_data$Immigration, levels = c("Strongly disagree","Disagree", "Neither","Agree","Strongly agree"))
```

The Original data is from survey by US GSS, which contains data of public opinions of USA citizens on various subjects. The data is from 2021 survey which was conducted online due to global pandemic by COVID19. The survey included a section asking people's opinion on various subjects regarding environment. To study the opinions of different groups of people about different aspects of environmental issues, the following variables from the survey has been chosen: age, gender, race,

One important thing that can be observed from the data, which is also important aspect of this research is is that there are many people who did not answer the questions regarding the environmental issues. This show that level of interest people show regarding this topic. Some people

There exists a survey similar to ours, CNN exit polls, however it is not easy to obtain the datasets they used. Also, US GSS is more realistic than exit polls since people would think more rational with sufficient time to think in US GSS than exit polls. Ext polls can show crucial factors, but it lackes the ability of general thoughts.

Now, let's get into the variables in details. One thing that we focused at first was whether respondents were equivalently recruited or not. To do this, you have to look at respondents' personal information. If you take a look from figure \@ref(fig:fig1) to \@ref(fig:fig4). You can see some interesting observations as followed:

1) Number of survey participants of age in the range of 20~40 is unexpectedly low compared to other ages.

2) There are significantly more white population compared other races. However, considering US race demographics, it is very reasonable. 
For more details on US race demographics, please visit the link: https://www.census.gov/quickfacts/fact/table/US/PST045221.

3) For gender, there is slightly more female than male, and we will see later whether this has impacts on vote or not.

From figure \@ref(fig:fig5) to \@ref(fig:fig14), they represent respondents' thoughts on national policies as well as relation between variables. Mostly they were collected from questions that whether you think that national spending on corresponding variables is about right, too much, or too little. For instance, let's take a look at figure \@ref(fig:fig5). This graph represents respondents' thoughts on U.S national spending on improving black condition, and we sorted it with our interest, Clinton or Trump. For more example, let's take a look at figure \@ref(fig:fig9).This graph represents relation between income and degree with respect to our interest. It generally shows that higher education you have, more income you earn. We are going to talk about interesting results in more detailed later in result and discussion session. For more information regarding variables graph please check Appendix.

```{r fig1,echo=FALSE, fig.cap = "Distribution of survey respondents' age group."}
reduced_data %>%
  drop_na(age_group) %>%
  ggplot(aes(x = age_group, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
  geom_bar() +
  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
  coord_flip() +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Survey Respondents' Age") +
  theme(legend.title = element_blank())
```

```{r fig2, fig.cap= "Distribution of survey respondents' race.", echo = FALSE}
reduced_data %>%
  ggplot(aes(x = race, fill = race)) +
  geom_bar() +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Survey Respondents' Race",
       ) +
  theme(legend.position = "none")
```

```{r fig3, echo = FALSE, fig.cap= "Distribution of survey respondents' gender"}
reduced_data %>%
  ggplot(aes(x = gender, fill = gender)) +
  geom_bar() +
  #scale_fill_manual("gender", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Survey Respondents' Gender",
       ) +
  theme(legend.position = "none")
```

```{r fig4, echo=FALSE, fig.cap = "Distribution of survey respondents"}
reduced_data %>%
  ggplot(aes(x = pres16, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
  geom_bar() +
  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Surveyee's Vote Choice of 2016 Presidential Election",
       ) +
  theme(legend.position = "none")
```

```{r fig5, echo=FALSE, fig.cap = "Opinion on national spending on improving conditions of black population"}
reduced_data %>%
  drop_na(nation_race) %>%
  drop_na(race) %>%
  ggplot(aes(x = race, fill = nation_race)) +
  geom_bar() +
  coord_flip() +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Is the US Spending Enough to Improve Conditions of Black Population?",
       ) +
  facet_wrap(~ pres16) + 
  theme(legend.title = element_blank())
```

```{r fig6, echo=FALSE,fig.cap = "Distribution of survey respondents' vote choice by birth place"}
reduced_data %>%
  ggplot(aes(x = born_US, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
  geom_bar() +
  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Surveyee's Vote Choice of 2016 Presidential Election by Brith Place",
       fil = "Candidate")
```

```{r fig7, echo=FALSE, fig.cap = "Distribution of survey respondents' vote by gender"}
reduced_data %>%
  ggplot(aes(x = pres16, fill = gender)) +
  geom_bar() +
  coord_flip() +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Surveyee's Vote Choice of 2016 Presidential Election by Gender"
       )
```

```{r fig8, echo=FALSE, fig.cap = "Opinion on national spending on welfare"}
reduced_data %>%
  drop_na(nation_welfare) %>%
  drop_na(race) %>%
  ggplot(aes(x = race, fill = nation_welfare)) +
  geom_bar() +
  coord_flip() +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Is the Government Spending Enough on Welfare?",
       ) +
  facet_wrap(~ pres16) + 
  theme(legend.title = element_blank())
```

```{r fig9, echo=FALSE, fig.cap = "Distribution of votes by degree and family income"}
reduced_data %>%
  drop_na(degree) %>%
  ggplot(aes(x = degree, fill = factor(Family_income, 
                                       levels = c("$170,000 or over","$130,000 ~ $169,999","$90,000 ~ $129,999","$60,000 ~ $89,999",
                                                  "$40,000 ~ $59,999","$20,000 ~ $39,999","$10,000 ~ $19,999","Under $10,000")))) +
  geom_bar(position = "dodge") +
  labs(fill = "Family Income") +
  coord_flip() +
  facet_wrap(~pres16) +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Education, Income and its Effect to the Vote"
       )
```

```{r fig10, echo=FALSE, fig.cap = "Opinion on wheter the US should limit immigration to protect their culture"}
reduced_data %>%
  ggplot(aes(x = Immigration, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
  geom_bar() +
  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
  labs(fil = "Candidate") +
  coord_flip() +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Should the Government Limit Immigration?"
       ) +
  facet_wrap(~pres16)
```

```{r fig11, echo=FALSE, fig.cap = "Surveyee's opinion on government's budget on education improvement"}
reduced_data %>%
  ggplot(aes(x = nation_education, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
  geom_bar(position = "dodge") +
  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
  labs(fil = "Candidate") +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Is the Government Spending Enough on Improving Education?"
       )
```

```{r fig12, echo=FALSE, fig.cap = "Surveyee's opinion on government's budget on arms"}
reduced_data %>%
  ggplot(aes(x = nation_arms, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
  geom_bar(position = "dodge") +
  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
  labs(fil = "Candidate") +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Is the Government Spending Enough on Arms and National Defense?"
       )
```

```{r fig13, echo=FALSE, fig.cap= "Surveyee's opinion on government's budget on assisting the poor"}
reduced_data %>%
  ggplot(aes(x = Assistance_to_the_poor, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
  geom_bar(position = "dodge") +
  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
  labs(fil = "Candidate") +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Is the Government Spending Enough to Assist the Poor?"
       )
```

```{r fig14, echo=FALSE, fig.cap = "Surveyee's opinion on government's budget on assisting other countries"}
reduced_data %>%
  ggplot(aes(x = Assistance_to_other_countries, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
  geom_bar(position = "dodge") +
  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
  labs(fil = "Candidate") +
  labs(x = element_blank(),
       y = element_blank(),
       title = "Is the Government Spending Enough to Assis Other Countries?"
       )
```

\newpage

# Results
The questions asking personal information shows some obvious but also interesting results. Figure \@ref(fig:fig1) shows that younger generation did not participate in the survey as much as older generation. It can also be seen that while almost all age groups preferred Clinton over Trump, compared to younger generation, older age groups show more interest to Trump. Females participants outnumbers male participants by significant amount as seen on Figure \@ref(fig:fig3), and many of female participants' vote went to Clinton, while male votes divided almost evenly(Figure \@ref(fig:fig6)). Figure \@ref(fig:fig8) shows some very interesting responses. Considering the distribution of votes, votes of white population are fairly split between Clinton and Trump, with more votes to Clinton, but all other races heavily favor Clinton. Some aspects that are not so obvious are ones involving education, family income and its correlation with the candidate preference. AS seen on figure \@ref(fig:fig9), People with higher education and higher income voted for Clinton while lower income and education votes are quite evenly split.

Figures 9-14 focuses on the questions that involves asking their opinions of various subjects of government policies and how supporters of different candidate, figure \@ref(fig:fig8) shows that white population's general opinion is that the government's budget on welfare is not lacking, with many of them thinking government is spending too much, whereas other races opinions are evenly split with black community leaning more towards the opinion that it is lacking. We can also observe from the same figure that Clinton supporters generally think that the budget on welfare is lacking, and Trump's supporters think budget is too much. Figure \@ref(fig:fig10) shows answers to question on immigration and whether the government should limit immigration to protect American culture, and it shows that out of the people who answered the question, majority of Clinton voters disagrees and majority of Trump voters agrees. We start to see a trend here that Trump supporters have more conservative opinions and Clinton supporters are more open to ideas on improvement in social equality or welfare. We see this trend continue on with other figures. On Figure \@ref(fig:fig12), many Clinton voters think government is spending too much of national defense while very few Trump voters think the same, while many trump voters think they spend too little when very few Clinton voters have the same opinion. For the question on if they think the government is spending enough to assist other countries (Figure \@ref(fig:fig14)), while Clinton's supporters opinions are divided quite evenly between "about right", and "too much", majority of Trump supporters think that the government is spending too much. Figure \@ref(fig:fig11) and Figure \@ref(fig:fig13) shows where Clinton voters and Trump voters share the same opinion on government spending on education and to assist the poor. While majority of people who thinks that the government is spending too much on both occasions are Trump supporters, popular opinions from both parties is that the government is spending too little on education and to assist the poor.

# Discussion
In this session, we are going to explore important questions that can be raised from our findings. At this point, you may end up with question marks in meanings this paper possess. Although this paper is a research of the past, this definitely has something to tell us about the U.S or even the world. 
Hopefully following questions helps you to understand our paper.

## Summary
We have taken survey data to observe the trend in voters of 2016 presidential election of the United States. In general, just as we expected by the trend we saw during our observation of the data, Hillary Clinton supporters were moved by improvement in social equality or welfare such as "Assistance to the Poor", "nation welfare", "nation education" and "Assistance to the other countries". On the other hand, Donald Trump supporters were interested by more conservative ideas that US is spending too much on the sessions above, and shows their support to Trump's slogan "Make America Great Again".

## Disagreement Between Races, Gender, Ages, and Wealth
Even though the United States and many other countries around the world have been trying hard to improve social welfare and equality, throughout this paper we have observed that there still exists disagreement within gender, race, wealth and age, and this had massive impact on politics and was crucial factor on people's decision on who to vote for as the president. Despite the popular movements voicing for equality between genders, and races and such in recent years, it seems that the opinions of people in different groups, are very divided.

## Weaknesses and how we should proceed in the future
The data from U.S GSS contained large amount of data and information and the distribution of survey respondents were quite accurate in terms of race. However, there were significantly more females than males and significantly less people of ages in the 20s, 30s and 40s. It is important, especially for the topic of this paper, to have the data on the younger generation and how they view government policies and how it effects their vote for the presidential election.
To gain more knowledge in this trend of disagreement between people of different race, gender, age and wealth when it comes to the election, an additional survey could be made to answer WHY this disagreement is happening, by not just asking for their opinion but also asking about what they experience and witness. In the future, survey should be including questions that can allow us to better capture and understand the current trend in agreement and disagreement between different groups of people, it can allow us to study how it could effect the election in the future.

\newpage

\appendix

# Appendix {-}
1) Figure \@ref(fig:fig6) represents people who were born in US and their votes on the candidates. Most respondents were born in US.

2) Figure \@ref(fig:fig8) represents national spending on welfare. People who advocated more funds in improvement of welfare shows preference in Clinton and people who did not shows preference in Trump.

3) Figure \@ref(fig:fig10) represents whether U.S should limit immigration in order to protect their own cultures or not. Even though many answered "Not Applicable", we can observe two different patters in Clinton and Trump.

4) Figure \@ref(fig:fig11) represents  national spending on education with votes. People who want more funds in education improvement shows preference in Clinton and vice versa.

5) Figure \@ref(fig:fig12) represents national spending on the arms with respect to votes for the candidates. People who thought too much clealry shows the preference toward Clinton.

6) Figure \@ref(fig:fig13) represents national spending on the poor with respect to respondents' votes on the candidates. From "About right" to "Too much", there exits more Trump supporters than Clinton's and vice versa.

7) Figure \@ref(fig:fig14) represents national spending on aids to foreign countries with respect to the votes for candidates . There exists more Trump supporters in "Too much" comparing to Clinton's supporters.

# Supplementary survey

Our supplementary survey is available here: https://docs.google.com/forms/d/e/1FAIpQLScN9ySPIn2W2DZ-Cs5GBpgXRcw79l1oNkqym3ARWWzIy-8uOg/viewform?usp=sf_link

## Preamble and Questions preview

The purpose of this survey is to further investigate the 2016 U.S presidential election. We are interested in more and specific opinions regarding factors you consider important to decide who to vote than what US GSS provided. If you are not a US citizen, please pretend you are, or imagine that having very similar candidates in your home country.

By proceeding with this survey, you understand that we will use your responses to 
better understand the 2016 U.S presidential election in the GSS. The survey is voluntary, and if you decide to participate you can skip questions and withdraw at any time.

## End Page

Thank you for participating in our survey. If you have any questions please do not hesitate to contact us.
sangwoong.lee@maill.utoronto.ca and young.suk@mail.utoronto.ca

\newpage

# References
Smith, Tom W., Davern, Michael, Freese, Jeremy, and Morgan, Stephen L., General Social Surveys, 1972-2018 [machine-readable data file] /Principal Investigator, Smith, Tom W.; Co-Principal Investigators, Michael Davern, Jeremy Freese and Stephen L. Morgan; Sponsored by National Science Foundation. --NORC ed.-- Chicago: NORC, 2019.

U.S. Census Bureau (2021). Population estimates. Retrieved from https://www.census.gov/quickfacts/fact/table/US/PST045221.

### 2022-04-20 scripts/01-data_cleaning.R

@@ -1,14 +1,11 @@
 #### Preamble ####
-# Purpose: Clean the survey data downloaded from [...UPDATE ME!!!!!]
-# Author: Rohan Alexander [CHANGE THIS TO YOUR NAME!!!!]
-# Data: 3 January 2021
-# Contact: rohan.alexander@utoronto.ca [PROBABLY CHANGE THIS ALSO!!!!]
+# Purpose: Prepare the 2021 US GSS data
+# Author: Young Suk
+# Data: 20 April 2022
+# Contact: young.suk@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
-# - Need to have downloaded the ACS data and saved it to inputs/data
-# - Don't forget to gitignore it!
-# - Change these to yours
-# Any other information needed?
+# - Need to have downloaded the GSS data and saved it to inputs/data
 
 
 #### Workspace setup ####
@@ -16,21 +13,175 @@
 library(haven)
 library(tidyverse)
 # Read in the raw data. 
-raw_data <- readr::read_csv("inputs/data/raw_data.csv"
-                     )
+raw_data <- haven::read_dta("inputs/data/2021_stata/gss2021.dta")
+
 # Just keep some variables that may be of interest (change 
 # this depending on your interests)
 names(raw_data)
 
 reduced_data <- 
   raw_data %>% 
-  select(first_col, 
-         second_col)
-rm(raw_data)
+  select(age,
+         sexnow1, 
+         degree,
+         degree,
+         income16,
+         raceacs1,
+         raceacs2,
+         raceacs3,
+         raceacs4,
+         raceacs5,
+         raceacs6,
+         raceacs7,
+         raceacs8,
+         raceacs9,
+         raceacs10,
+         raceacs15,
+         raceacs16,
+         nateduc,
+         grncon,
+         enprbus,
+         clmtcaus,
+         clmtwrld,
+         scigrn,
+         grnecon,
+         grnprog,
+         grntaxes,
+         naturdev
          
+  ) %>% 
+  rename(gender = sexnow1,
+         Family_income = income16,
+         nation_education = nateduc)
 
-#### What's next? ####
+#rm(raw_data)
+
+# Add age_group column based on age column
+reduced_data <- reduced_data %>% 
+  mutate(age_group = reduced_data$age %>% cut(breaks = seq(19, 89, by = 10))) %>%
[…]

### 2022-04-22 outputs/paper/paper.Rmd
data cleaning, ggplot

@@
 # Introduction
-Environmental issues around the world as been very important part of modern society as many changes in the environment such as climate change, water and air pollution started to be very noticeable.
-While the concerns regarding the state of the environment has become popular issue, it is still is not
+Religion has been major part of human history around the world, and for many countries, religion is something they heavily rely on even in modern days.For example, European countries such as Spain and Italy which heavily relies on catholic, and there are many middle eastern countries that heavily rely on muslim, islam and hinduism.
+While it is reported that the number of religious people is declining quickly, religion still has big social influence in many aspects of the society in the United States, and this paper aims to research the social influence of popular religions in the United States and how the opinions of different religions differ in the context of social issues and government policies regarding them.
 
-Donald Trump won the 2016 United States presidential election.
-From what we remembered, it was quite surprising considering his image outside of U.S or even in U.S.
-Also, before the election, Hillary Clinton was ahead in most surveys as well.
-There is a famous quote that we learn from the past.
-We thought that even though we are living in 2022, it would be meaningful to analyze the election so that we get to know what U.S citizens thought, or in broader sense, thoughts trends in nowadays and what factors we considered most important to decide our representative.
+This research will focus on some social issues in the United States and how opinions of people of different religion differs.
 
@@
 reduced_data <- read_csv("outputs/data/prepared_gss.csv")
-reduced_data$pres16 <- factor(reduced_data$pres16, levels = c("Clinton","Trump","Other","Did not vote"))
 # Change the order of variables for graphing purposes
@@
 The data is from 2021 survey which was conducted online due to global pandemic by COVID19.
-The survey included a section asking people's opinion on various subjects regarding environment.
-To study the opinions of different groups of people about different aspects of environmental issues, the following variables from the survey has been chosen: age, gender, race,
+For the purpose of this research, the data related to personal profiles, survey respondants' opinion on government policies and religion has been extracted from the raw survey data.
 
-One important thing that can be observed from the data, which is also important aspect of this research is is that there are many people who did not answer the questions regarding the environmental issues.
-This show that level of interest people show regarding this topic.
-Some people
-
-There exists a survey similar to ours, CNN exit polls, however it is not easy to obtain the datasets they used.
-Also, US GSS is more realistic than exit polls since people would think more rational with sufficient time to think in US GSS than exit polls.
-Ext polls can show crucial factors, but it lackes the ability of general thoughts.
-
-Now, let's get into the variables in details.
-One thing that we focused at first was whether respondents were equivalently recruited or not.
-To do this, you have to look at respondents' personal information.
-If you take a look from figure \@ref(fig:fig1) to \@ref(fig:fig4).
-You can see some interesting observations as followed:
-
-1) Number of survey participants of age in the range of 20~40 is unexpectedly low compared to other ages.
-
-2) There are significantly more white population compared other races. However, considering US race demographics, it is very reasonable. 
-For more details on US race demographics, please visit the link: https://www.census.gov/quickfacts/fact/table/US/PST045221.
-
-3) For gender, there is slightly more female than male, and we will see later whether this has impacts on vote or not.
-
-From figure \@ref(fig:fig5) to \@ref(fig:fig14), they represent respondents' thoughts on national policies as well as relation between variables.
-Mostly they were collected from questions that whether you think that national spending on corresponding variables is about right, too much, or too little.
-For instance, let's take a look at figure \@ref(fig:fig5).
-This graph represents respondents' thoughts on U.S national spending on improving black condition, and we sorted it with our interest, Clinton or Trump.
-For more example, let's take a look at figure \@ref(fig:fig9).This graph represents relation between income and degree with respect to our interest.
-It generally shows that higher education you have, more income you earn.
-We are going to talk about interesting results in more detailed later in result and discussion session.
-For more information regarding variables graph please check Appendix.
+The religion data has very detailed options that includes religions have very low For the simplicity of this research, protestants and christians were grouped together under chrinstans and all the non-popular religion has been grouped under "other".
 
@@
   drop_na(age_group) %>%
-  ggplot(aes(x = age_group, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  ggplot(aes(x = age_group, fill = relig)) +
   geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
   coord_flip() +
@@
 reduced_data %>%
-  ggplot(aes(x = race, fill = race)) +
+  ggplot(aes(x = race, fill = relig)) +
   geom_bar() +
@@
 reduced_data %>%
-  ggplot(aes(x = gender, fill = gender)) +
+  ggplot(aes(x = gender, fill = relig)) +
   geom_bar() +
-  #scale_fill_manual("gender", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
   labs(x = element_blank(),
@@
 
-```{r fig4, echo=FALSE, fig.cap = "Distribution of survey respondents"}
+```{r fig4, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
 reduced_data %>%
-  ggplot(aes(x = pres16, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
+  ggplot(aes(x = relig, fill = relig)) +
   geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
   labs(x = element_blank(),
        y = element_blank(),
-       title = "Surveyee's Vote Choice of 2016 Presidential Election",
+       title = "Survey Respondents' Religion",
        ) +
-  theme(legend.position = "none")
-```
-
-```{r fig5, echo=FALSE, fig.cap = "Opinion on national spending on improving conditions of black population"}
-reduced_data %>%
-  drop_na(nation_race) %>%
-  drop_na(race) %>%
-  ggplot(aes(x = race, fill = nation_race)) +
-  geom_bar() +
   coord_flip() +
-  labs(x = element_blank(),
-       y = element_blank(),
-       title = "Is the US Spending Enough to Improve Conditions of Black Population?",
-       ) +
-  facet_wrap(~ pres16) + 
   theme(legend.title = element_blank())
-```
-
-```{r fig6, echo=FALSE,fig.cap = "Distribution of survey respondents' vote choice by birth place"}
-reduced_data %>%
-  ggplot(aes(x = born_US, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-  labs(x = element_blank(),
-       y = element_blank(),
-       title = "Surveyee's Vote Choice of 2016 Presidential Election by Brith Place",
-       fil = "Candidate")
-```
-
-```{r fig7, echo=FALSE, fig.cap = "Distribution of survey respondents' vote by gender"}
-reduced_data %>%
-  ggplot(aes(x = pres16, fill = gender)) +
-  geom_bar() +
-  coord_flip() +
-  labs(x = element_blank(),
-       y = element_blank(),
-       title = "Surveyee's Vote Choice of 2016 Presidential Election by Gender"
-       )
-```
-
-```{r fig8, echo=FALSE, fig.cap = "Opinion on national spending on welfare"}
-reduced_data %>%
-  drop_na(nation_welfare) %>%
-  drop_na(race) %>%
-  ggplot(aes(x = race, fill = nation_welfare)) +
-  geom_bar() +
-  coord_flip() +
-  labs(x = element_blank(),
-       y = element_blank(),
-       title = "Is the Government Spending Enough on Welfare?",
-       ) +
-  facet_wrap(~ pres16) + 
-  theme(legend.title = element_blank())
-```
-
-```{r fig9, echo=FALSE, fig.cap = "Distribution of votes by degree and family income"}
-reduced_data %>%
-  drop_na(degree) %>%
-  ggplot(aes(x = degree, fill = factor(Family_income, 
-                                       levels = c("$170,000 or over","$130,000 ~ $169,999","$90,000 ~ $129,999","$60,000 ~ $89,999",
-                                                  "$40,000 ~ $59,999","$20,000 ~ $39,999","$10,000 ~ $19,999","Under $10,000")))) +
-  geom_bar(position = "dodge") +
-  labs(fill = "Family Income") +
-  coord_flip() +
-  facet_wrap(~pres16) +
-  labs(x = element_blank(),
-       y = element_blank(),
-       title = "Education, Income and its Effect to the Vote"
-       )
-```
-
-```{r fig10, echo=FALSE, fig.cap = "Opinion on wheter the US should limit immigration to protect their culture"}
-reduced_data %>%
-  ggplot(aes(x = Immigration, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar() +
-  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-  labs(fil = "Candidate") +
-  coord_flip() +
-  labs(x = element_blank(),
-       y = element_blank(),
-       title = "Should the Government Limit Immigration?"
-       ) +
-  facet_wrap(~pres16)
-```
-
-```{r fig11, echo=FALSE, fig.cap = "Surveyee's opinion on government's budget on education improvement"}
-reduced_data %>%
-  ggplot(aes(x = nation_education, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar(position = "dodge") +
-  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-  labs(fil = "Candidate") +
-  labs(x = element_blank(),
-       y = element_blank(),
-       title = "Is the Government Spending Enough on Improving Education?"
-       )
-```
-
-```{r fig12, echo=FALSE, fig.cap = "Surveyee's opinion on government's budget on arms"}
-reduced_data %>%
-  ggplot(aes(x = nation_arms, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar(position = "dodge") +
-  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-  labs(fil = "Candidate") +
-  labs(x = element_blank(),
-       y = element_blank(),
-       title = "Is the Government Spending Enough on Arms and National Defense?"
-       )
-```
-
-```{r fig13, echo=FALSE, fig.cap= "Surveyee's opinion on government's budget on assisting the poor"}
-reduced_data %>%
-  ggplot(aes(x = Assistance_to_the_poor, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar(position = "dodge") +
-  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-  labs(fil = "Candidate") +
-  labs(x = element_blank(),
-       y = element_blank(),
-       title = "Is the Government Spending Enough to Assist the Poor?"
-       )
-```
-
-```{r fig14, echo=FALSE, fig.cap = "Surveyee's opinion on government's budget on assisting other countries"}
-reduced_data %>%
-  ggplot(aes(x = Assistance_to_other_countries, fill = factor(pres16, levels = c("Clinton", "Trump", "Other", "Did not vote")))) +
-  geom_bar(position = "dodge") +
-  scale_fill_manual("Candidate", values = c("Clinton" = "#01378C", "Trump" = "#ff3d3d", "Other" = "#2cd18c", "Did not vote" = "#dede1b")) +
-  labs(fil = "Candidate") +
-  labs(x = element_blank(),
-       y = element_blank(),
-       title = "Is the Government Spending Enough to Assis Other Countries?"
-       )
 ```
@@
 # Results
-The questions asking personal information shows some obvious but also interesting results.
-Figure \@ref(fig:fig1) shows that younger generation did not participate in the survey as much as older generation.
-It can also be seen that while almost all age groups preferred Clinton over Trump, compared to younger generation, older age groups show more interest to Trump.
-Females participants outnumbers male participants by significant amount as seen on Figure \@ref(fig:fig3), and many of female participants' vote went to Clinton, while male votes divided almost evenly(Figure \@ref(fig:fig6)).
-Figure \@ref(fig:fig8) shows some very interesting responses.
-Considering the distribution of votes, votes of white population are fairly split between Clinton and Trump, with more votes to Clinton, but all other races heavily favor Clinton.
-Some aspects that are not so obvious are ones involving education, family income and its correlation with the candidate preference.
-AS seen on figure \@ref(fig:fig9), People with higher education and higher income voted for Clinton while lower income and education votes are quite evenly split.
-
-Figures 9-14 focuses on the questions that involves asking their opinions of various subjects of government policies and how supporters of different candidate, figure \@ref(fig:fig8) shows that white population's general opinion is that the government's budget on welfare is not lacking, with many of them thinking government is spending too much, whereas other races opinions are evenly split with black community leaning more towards the opinion that it is lacking.
-We can also observe from the same figure that Clinton supporters generally think that the budget on welfare is lacking, and Trump's supporters think budget is too much.
-Figure \@ref(fig:fig10) shows answers to question on immigration and whether the government should limit immigration to protect American culture, and it shows that out of the people who answered the question, majority of Clinton voters disagrees and majority of Trump voters agrees.
-We start to see a trend here that Trump supporters have more conservative opinions and Clinton supporters are more open to ideas on improvement in social equality or welfare.
-We see this trend continue on with other figures.
-On Figure \@ref(fig:fig12), many Clinton voters think government is spending too much of national defense while very few Trump voters think the same, while many trump voters think they spend too little when very few Clinton voters have the same opinion.
-For the question on if they think the government is spending enough to assist other countries (Figure \@ref(fig:fig14)), while Clinton's supporters opinions are divided quite evenly between "about right", and "too much", majority of Trump supporters think that the government is spending too much.
-Figure \@ref(fig:fig11) and Figure \@ref(fig:fig13) shows where Clinton voters and Trump voters share the same opinion on government spending on education and to assist the poor.
-While majority of people who thinks that the government is spending too much on both occasions are Trump supporters, popular opinions from both parties is that the government is spending too little on education and to assist the poor.
 
 # Discussion
-In this session, we are going to explore important questions that can be raised from our findings.
-At this point, you may end up with question marks in meanings this paper possess.
-Although this paper is a research of the past, this definitely has something to tell us about the U.S or even the world.
-Hopefully following questions helps you to understand our paper.
 
 ## Summary
-We have taken survey data to observe the trend in voters of 2016 presidential election of the United States.
-In general, just as we expected by the trend we saw during our observation of the data, Hillary Clinton supporters were moved by improvement in social equality or welfare such as "Assistance to the Poor", "nation welfare", "nation education" and "Assistance to the other countries".
-On the other hand, Donald Trump supporters were interested by more conservative ideas that US is spending too much on the sessions above, and shows their support to Trump's slogan "Make America Great Again".
 
-## Disagreement Between Races, Gender, Ages, and Wealth
-Even though the United States and many other countries around the world have been trying hard to improve social welfare and equality, throughout this paper we have observed that there still exists disagreement within gender, race, wealth and age, and this had massive impact on politics and was crucial factor on people's decision on who to vote for as the president.
-Despite the popular movements voicing for equality between genders, and races and such in recent years, it seems that the opinions of people in different groups, are very divided.
+## Different Opinions between different religions.
 
 ## Weaknesses and how we should proceed in the future
-The data from U.S GSS contained large amount of data and information and the distribution of survey respondents were quite accurate in terms of race.
-However, there were significantly more females than males and significantly less people of ages in the 20s, 30s and 40s.
-It is important, especially for the topic of this paper, to have the data on the younger generation and how they view government policies and how it effects their vote for the presidential election.
-To gain more knowledge in this trend of disagreement between people of different race, gender, age and wealth when it comes to the election, an additional survey could be made to answer WHY this disagreement is happening, by not just asking for their opinion but also asking about what they experience and witness.
-In the future, survey should be including questions that can allow us to better capture and understand the current trend in agreement and disagreement between different groups of people, it can allow us to study how it could effect the election in the future.
 
@@
 
-\appendix
-
-# Appendix {-}
-1) Figure \@ref(fig:fig6) represents people who were born in US and their votes on the candidates. Most respondents were born in US.
-
-2) Figure \@ref(fig:fig8) represents national spending on welfare.
-People who advocated more funds in improvement of welfare shows preference in Clinton and people who did not shows preference in Trump.
-
-3) Figure \@ref(fig:fig10) represents whether U.S should limit immigration in order to protect their own cultures or not.
-Even though many answered "Not Applicable", we can observe two different patters in Clinton and Trump.
-
-4) Figure \@ref(fig:fig11) represents  national spending on education with votes. People who want more funds in education improvement shows preference in Clinton and vice versa.
-
-5) Figure \@ref(fig:fig12) represents national spending on the arms with respect to votes for the candidates. People who thought too much clealry shows the preference toward Clinton.
-
-6) Figure \@ref(fig:fig13) represents national spending on the poor with respect to respondents' votes on the candidates.
-From "About right" to "Too much", there exits more Trump supporters than Clinton's and vice versa.
-
-7) Figure \@ref(fig:fig14) represents national spending on aids to foreign countries with respect to the votes for candidates .
-There exists more Trump supporters in "Too much" comparing to Clinton's supporters.
-
-# Supplementary survey
-
-Our supplementary survey is available here: https://docs.google.com/forms/d/e/1FAIpQLScN9ySPIn2W2DZ-Cs5GBpgXRcw79l1oNkqym3ARWWzIy-8uOg/viewform?usp=sf_link
-
-## Preamble and Questions preview
-
-The purpose of this survey is to further investigate the 2016 U.S presidential election.
-We are interested in more and specific opinions regarding factors you consider important to decide who to vote than what US GSS provided.
-If you are not a US citizen, please pretend you are, or imagine that having very similar candidates in your home country.
-
-By proceeding with this survey, you understand that we will use your responses to 
-better understand the 2016 U.S presidential election in the GSS. The survey is voluntary, and if you decide to participate you can skip questions and withdraw at any time.
-
-## End Page
-
-Thank you for participating in our survey. If you have any questions please do not hesitate to contact us.
-sangwoong.lee@maill.utoronto.ca and young.suk@mail.utoronto.ca
+# appendix
 

### 2022-04-22 scripts/01-data_cleaning.R

@@ -24,7 +24,6 @@ reduced_data <-
   select(age,
          sexnow1, 
          degree,
-         degree,
          income16,
          raceacs1,
          raceacs2,
@@ -38,23 +37,27 @@ reduced_data <-
          raceacs10,
          raceacs15,
          raceacs16,
-         nateduc,
-         grncon,
-         enprbus,
-         clmtcaus,
-         clmtwrld,
-         scigrn,
-         grnecon,
-         grnprog,
-         grntaxes,
-         naturdev
+         natrace,
+         natfare,
+         natfarey,
+         natsoc,
+         natcrime,
+         immlimit,
+         relig,
+         religinf,
+         opsex,
          
   ) %>% 
   rename(gender = sexnow1,
          Family_income = income16,
-         nation_education = nateduc)
+         nation_race = natrace,
+         nation_welfare = natfare,
+         Assistance_to_the_poor = natfarey,
+         social_security = natsoc,
+         Immigration = immlimit,
+         )
 
-#rm(raw_data)
+# rm(raw_data)
 
 # Add age_group column based on age column
 reduced_data <- reduced_data %>% 
@@ -69,7 +72,17 @@ reduced_data <- reduced_data %>%
     age_group == "(79,89]" ~ "80-89 years old"
   ))
 
-
+# # Recode to rename variables in imigration according to options from the Codebook 'GSS 2021 Codebook R1b.pdf'.
+# The question was America should limit immigration in order to protect our national way of life
+reduced_data <- 
+  reduced_data %>% 
+  mutate(Immigration = case_when(
+    Immigration == 1 ~ "Strongly agree",
+    Immigration == 2 ~ "Agree",
+    Immigration == 3 ~ "Neither",
+    Immigration == 4 ~ "Disagree",
+    Immigration == 5 ~ "Strongly disagree"
+  ))
 # Recode to rename variables in gender according to options from the Codebook 'GSS 2021 Codebook R1b.pdf'.
 # The question was Do you describe yourself as male, female, or transgender?
 reduced_data <- 
@@ -80,6 +93,7 @@ reduced_data <-
     gender == 3 ~ "Transgender",
     gender == 4 ~ "Other",
   ))
+
 
 # Recode to rename variables in degree according to options from the Codebook 'GSS 2021 Codebook R1b.pdf'.
 # The question was the respondent's degree.
@@ -92,6 +106,7 @@ reduced_data <-
[…]

### 2022-04-23 outputs/paper/paper.Rmd
ggplot

@@
 ---
-title: "Disagreement Regarding 2016 Presidential Election in the United States^[Code and data are available at: https://github.com/SANGWOONG-LEE/2016electionanalysis]"
+title: "Religion and Its Influence on Social Issues^[Code and data are available at: https://github.com/dudqo/religion_survey]"
 subtitle: "Possible factors effecting 2016 election"
@@
 The data is from 2021 survey which was conducted online due to global pandemic by COVID19.
-For the purpose of this research, the data related to personal profiles, survey respondants' opinion on government policies and religion has been extracted from the raw survey data.
+For the purpose of this research, the data related to personal profiles, survey respondents' opinion on government policies and religion has been extracted from the raw survey data.
 
-The religion data has very detailed options that includes religions have very low For the simplicity of this research, protestants and christians were grouped together under chrinstans and all the non-popular religion has been grouped under "other".
+The religion data has very detailed options that includes religions have very low For the simplicity of this research, protestants and Christians were grouped together under Christians and all the non-popular religion has been grouped under "other".
+
+Looking at the figures, some expected trends from the data can be observed:
+
+* As age gets younger, the percentage of people who does not have any religion increases.
+With steady increase in nonreligious as age group gets younger, people in the age group of 80-89 have very few people who are not religious.
+If we look at 20-29 years old, we can see that about half of them does not have any religion.
+* There are far more white population who participated in the survey than any other race. This is as expected since white population in the United States far outnumbers any other race.
+* Christian is most popular religion in the United States.
 
@@
 reduced_data %>%
+  ggplot(aes(x = relig, fill = Family_income)) +
+  geom_bar() +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Survey Respondents' Religion",
+       ) +
+  coord_flip() +
+  theme(legend.title = element_blank())
+```
+```{r fig5, echo=FALSE, fig.cap = "Distribution of religion of survey respondents by family income"}
+reduced_data %>%
   ggplot(aes(x = relig, fill = relig)) +
@@
        y = element_blank(),
-       title = "Survey Respondents' Religion",
+       title = "Survey Respondents' Family Income by Religion",
+       ) +
+  coord_flip() +
+  theme(legend.title = element_blank())
+```
+
+```{r fig6, echo=FALSE, fig.cap = "Survey respondents' opinion on governmemt policy on improving coditions of black population"}
+reduced_data %>%
+  drop_na(nation_race) %>%
+  ggplot(aes(x = relig, fill = nation_race)) +
+  geom_bar() +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Is the Government spending enough on improving coditions of Black People?",
+       ) +
+  coord_flip() +
+  theme(legend.title = element_blank())
+```
+```{r fig7, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+reduced_data %>%
+  drop_na(nation_welfare) %>%
+  ggplot(aes(x = relig, fill = nation_welfare)) +
+  geom_bar() +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Is the Government spending enough on coditions of Black People?",
+       ) +
+  coord_flip() +
+  theme(legend.title = element_blank())
+```
+
+```{r fig8, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+reduced_data %>%
+  drop_na(Assistance_to_the_poor) %>%
+  ggplot(aes(x = relig, fill = Assistance_to_the_poor)) +
+  geom_bar() +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Is the Government spending enough on coditions of Black People?",
        ) +

### 2022-04-23 scripts/01-data_cleaning.R

@@ -213,7 +213,7 @@ reduced_data <- reduced_data %>%
 
 # Remove columns that are no longer needed.
 reduced_data <- reduced_data %>%
-  select(-c(raceacs1, raceacs2, raceacs3, raceacs4, raceacs5, raceacs6, raceacs7, raceacs8, raceacs9, raceacs10, raceacs15, raceacs16, race_sum))
+  select(-c(age, raceacs1, raceacs2, raceacs3, raceacs4, raceacs5, raceacs6, raceacs7, raceacs8, raceacs9, raceacs10, raceacs15, raceacs16, race_sum))
 
 
 

### 2022-04-23 outputs/paper/paper.Rmd
ggplot

@@
        ) +
-  theme(legend.position = "none")
+  theme(legend.title = element_blank()) +
+  coord_flip() +
+  facet_wrap(~opsex)
 ```
@@
        y = element_blank(),
-       title = "Is the Government spending enough on coditions of Black People?",
+       title = "Is the Government Spending Enough to Asssist the Poor?",
+       ) +
+  coord_flip() +
+  theme(legend.title = element_blank())
+```
+```{r fig9, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+reduced_data %>%
+  drop_na(social_security) %>%
+  ggplot(aes(x = relig, fill = social_security)) +
+  geom_bar() +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Is the Government Spending Enough to Asssist the Poor?",
+       ) +
+  coord_flip() +
+  theme(legend.title = element_blank())
+```
+```{r fig10, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+reduced_data %>%
+  drop_na(natcrime) %>%
+  ggplot(aes(x = relig, fill = natcrime)) +
+  geom_bar() +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Is the Government Spending Enough to Asssist the Poor?",
+       ) +
+  coord_flip() +
+  theme(legend.title = element_blank())
+```
+```{r fig11, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+reduced_data %>%
+  drop_na(Immigration) %>%
+  ggplot(aes(x = relig, fill = Immigration)) +
+  geom_bar() +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Is the Government Spending Enough to Asssist the Poor?",
+       ) +
+  coord_flip() +
+  theme(legend.title = element_blank())
+```
+
+```{r fig12, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+reduced_data %>%
+  drop_na(religinf) %>%
+  ggplot(aes(x = relig, fill = religinf)) +
+  geom_bar() +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Is the Government Spending Enough to Asssist the Poor?",
+       ) +
+  coord_flip() +
+  theme(legend.title = element_blank())
+```
+```{r fig13, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+reduced_data %>%
+  drop_na(opsex) %>%
+  ggplot(aes(x = relig, fill = opsex)) +
+  geom_bar() +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Is the Government Spending Enough to Asssist the Poor?",
        ) +

### 2022-04-24 outputs/paper/paper.Rmd
reference and introduction

--- 2022-04-23 outputs/paper/paper.Rmd
@@
 # Introduction
-Religion has been major part of human history around the world, and for many countries, religion is something they heavily rely on even in modern days.For example, European countries such as Spain and Italy which heavily relies on catholic, and there are many middle eastern countries that heavily rely on muslim, islam and hinduism.
+Religion has been major part of human history around the world, and for many countries, religion is something they heavily rely on even in modern days.For example, European countries such as Spain and Italy which heavily relies on catholic, and there are many middle eastern countries that heavily rely on Islam and Hinduism.
 While it is reported that the number of religious people is declining quickly, religion still has big social influence in many aspects of the society in the United States, and this paper aims to research the social influence of popular religions in the United States and how the opinions of different religions differ in the context of social issues and government policies regarding them.
@@
 
-This paper is mainly going to explore possible variables that potentially had impacts on decision regarding the election as well as their analysis.
-However, making a choice for  which variables affected U.S citizens to lean on Trump side or Clinton side was not easy because there is nothing in this world absolute.
-Thus, as we mentioned above, we are going to focus on general flows of results instead of defining variables as absolute factors of the demarcation.
-Following that, this paper has structures: Data, Results, Discussion, Appendix and supplementary survey.
-
-The dataset was obtained from U.S GSS, and only picked and used the variables that are very related to the subject of this study.
-Throughout the analysis of relation between the variables and respondents' votes on the election, we could observe that people who had conservative thoughts on equality and welfare preferred Trump, and people who did not preferred Clinton.
-This may sound obvious considering their parties.
-In the result session, reader can observe that what specific areas or fields the respondents considered important.
-We believe that this situation is not only happening in U.S, but it is also happening all over the world.
-This claim is going to be conducted with supplementary survey session.
+The dataset was obtained from U.S GSS survey, by only selecting the survey questions that are related to the topic of this research.
+The data used includes personal profile data of survey respondents such as age, gender, religion, race, income and data on questions regarding their opinion on popular social issues in the United States such as abortion, gay marriage and government policies regarding welfare, racism, and etc.
+This research found that despite the people's different views on different religion and the differences in the religions themselves, when it comes to the popular social issues, their opinions are mostly inline with each other.
+Some different views do exists and which will be discussed in the later part of the paper.
 
@@
        y = element_blank(),
-       title = "Survey Respondents' Gender",
+       title = "Gender and Opinion on Importance of Being Born as Male or Female",
        ) +
-  theme(legend.position = "none")
+  theme(legend.title = element_blank()) +
+  coord_flip() +
+  facet_wrap(~opsex)
+```
+```{r fig4, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+reduced_data %>%
+  ggplot(aes(x = relig, fill = born)) +
+  geom_bar() +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Number of Survey Respondents' Born in U.S",
+       ) +
+  coord_flip() +
+  theme(legend.title = element_blank())
 ```
 
-```{r fig4, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+```{r fig5, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
 reduced_data %>%
@@
 ```
-```{r fig5, echo=FALSE, fig.cap = "Distribution of religion of survey respondents by family income"}
+```{r fig6, echo=FALSE, fig.cap = "Distribution of religion of survey respondents by family income"}
 reduced_data %>%
@@
 
-```{r fig6, echo=FALSE, fig.cap = "Survey respondents' opinion on governmemt policy on improving coditions of black population"}
+```{r fig7, echo=FALSE, fig.cap = "Survey respondents' opinion on governmemt policy on improving coditions of black population"}
 reduced_data %>%
@@
 ```
-```{r fig7, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+```{r fig8, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
 reduced_data %>%
@@
        y = element_blank(),
-       title = "Is the Government spending enough on coditions of Black People?",
+       title = "Is the Government Spending Enough on Welfare?",
        ) +
@@
 
-```{r fig8, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+```{r fig9, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
 reduced_data %>%
@@
        y = element_blank(),
-       title = "Is the Government spending enough on coditions of Black People?",
+       title = "Is the Government Spending Enough to Asssist the Poor?",
+       ) +
+  coord_flip() +
+  theme(legend.title = element_blank())
+```
+```{r fig10, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+reduced_data %>%
+  drop_na(social_security) %>%
+  ggplot(aes(x = relig, fill = social_security)) +
+  geom_bar() +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Is the Government Spending Enough on Social Security?",
+       ) +
+  coord_flip() +
+  theme(legend.title = element_blank())
+```
+```{r fig11, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+reduced_data %>%
+  drop_na(natcrime) %>%
+  ggplot(aes(x = relig, fill = natcrime)) +
+  geom_bar() +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Is the Government Spending Enough to Halt Increasing Crime Rate?",
+       ) +
+  coord_flip() +
+  theme(legend.title = element_blank())
+```
+```{r fig12, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+reduced_data %>%
+  drop_na(Immigration) %>%
+  ggplot(aes(x = relig, fill = Immigration)) +
+  geom_bar() +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Should the Government Limit Immigration to Protect U.S Culture?",
+       ) +
+  coord_flip() +
+  theme(legend.title = element_blank())
+```
+
+```{r fig13, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+reduced_data %>%
+  drop_na(religinf) %>%
+  ggplot(aes(x = relig, fill = religinf)) +
+  geom_bar() +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Would the U.S be a better country if religion had less influence?",
+       ) +
+  coord_flip() +
+  theme(legend.title = element_blank())
+```
+```{r fig14, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+reduced_data %>%
+  drop_na(homo_marriage) %>%
+  ggplot(aes(x = relig, fill = homo_marriage)) +
+  geom_bar() +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Do You Agree With Same Sex Marriage?",
+       ) +
+  coord_flip() +
+  theme(legend.title = element_blank())
+```
+```{r fig15, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+reduced_data %>%
+  drop_na(abortion) %>%
+  ggplot(aes(x = relig, fill = abortion)) +
+  geom_bar() +
+  labs(x = element_blank(),
+       y = element_blank(),
+       title = "Do you agree with abortion for any reason?",
        ) +
@@
 # Results
+To study more on gender and the opinions on importance of being born as certain gender, linear model was created to study this matter.
+
+As for the opinions on abortion, we can see an interesting result where only Christians and Catholics show significant amount of negativity towards abortion compared to other religions.
 
@@
 ## Summary
+In american society, general public's opinions on different religion varies significantly.
+While religions such as Christianity and Catholicism has been around since the beginning of history of the United States as a country, there are some religions such as Islam and Buddhism that Americans still are not very familiar with.
+With the unfamiliarity and some negative news involving a religion, the views on a specific religion be heavily influenced by stereotypes.
+We can observe from the u.S GSS survey that the opinions between people of different religion are not that different.
 
@@
 ## Weaknesses and how we should proceed in the future
+As the survey was conducted in the United States, the data lacks diversity of religions and most of survey respondents' are Christians or Catholics.
+For more accurate and in depth research, more data from people of other religions are required.
+Also, the opinions regarding social issues could be influenced by the location they live.
+So data from different countries where Christianity or Catholicism is not the main religion could be very helpful for this research.
 
@@
 
-# appendix
+# Appendix
 
@@
 # References
-Smith, Tom W., Davern, Michael, Freese, Jeremy, and Morgan, Stephen L., General Social Surveys, 1972-2018 [machine-readable data file] /Principal Investigator, Smith, Tom W.; Co-Principal Investigators, Michael Davern, Jeremy Freese and Stephen L.
-Morgan; Sponsored by National Science Foundation.
---NORC ed.-- Chicago: NORC, 2019.
-
-U.S. Census Bureau (2021). Population estimates. Retrieved from https://www.census.gov/quickfacts/fact/table/US/PST045221.

### 2022-04-24 scripts/01-data_cleaning.R

@@ -24,6 +24,7 @@ reduced_data <-
   select(age,
          sexnow1, 
          degree,
+         born,
          income16,
          raceacs1,
          raceacs2,
@@ -42,10 +43,13 @@ reduced_data <-
          natfarey,
          natsoc,
          natcrime,
+         natdrug,
          immlimit,
          relig,
          religinf,
          opsex,
+         marhomo,
+         abany
          
   ) %>% 
   rename(gender = sexnow1,
@@ -55,6 +59,8 @@ reduced_data <-
          Assistance_to_the_poor = natfarey,
          social_security = natsoc,
          Immigration = immlimit,
+         abortion = abany,
+         homo_marriage = marhomo
          )
 
 # rm(raw_data)
@@ -180,9 +186,11 @@ reduced_data <-
     relig == 9 ~ "Muslim/Islam",
     relig == 10 ~ "Other",
     relig == 11 ~ "Christian",
-    relig == 12 ~ "Native American",
+    relig == 12 ~ "Other",
     relig == 13 ~ "Other",
   ))
+
+
 
 # First step of merging process of survey response related to race. Create a column race_sum which each row represents number of races of each person.
 # If value at column race_sum is greater than 1, that person is of mixed race.
@@ -217,10 +225,9 @@ reduced_data <- reduced_data %>%
 
 
 
-# Recode to rename variables in nation_welfare according to options from the Codebook 'GSS 2021 Codebook R1b.pdf'.
-# The question was whether US is spending too much, too little, or about the right amount on 
-# Welfare
-# Unfortunetly 50.7% of respondents answered "Not Applicable" which falls into "NA".
+
+# Rename values of variable nation_welfare using codebook 'GSS 2021 Codebook R1b.pdf'.
+# survey question: Is the government spending enough on welfare?
 reduced_data <- 
   reduced_data %>% 
   mutate(nation_welfare = case_when(
@@ -229,11 +236,9 @@ reduced_data <-
     nation_welfare == 3 ~ "Too much",
   ))
 
-# Recode to rename variables in Assistance_to_the_poor according to options from the Codebook 'GSS 2021 Codebook R1b.pdf'.
-# The question was whether US is spending too much, too little, or about the right amount on 
-# assistance to the poor
-# Unfortunetly 49.3% of respondents answered "Not Applicable" which falls into "NA".
 
[…]

### 2022-04-27 outputs/paper/paper.Rmd
model

@@
 library(knitr)
+library(modelsummary)
 ```
@@
 # Results
-To study more on gender and the opinions on importance of being born as certain gender, linear model was created to study this matter.
+To study more on the social issues and the influence of religion, we are going to create some linear models.
+As opinions of different religions are similar in many of the survey question studied, only the questions that resulted in some disagreement will be chosen to create the models.
+Those data are abortion, same sex marriage, and opinion on religion influence.
+As for the predictors, age, gender, race, religion are used to fit the model.
+
+```{r}
+reduced_data <- reduced_data %>%
+  drop_na(abortion) %>%
+  drop_na(homo_marriage) %>%
+  drop_na(religinf)
+
+abortion_model <- glm(abortion ~ age_group + gender + race + relig, data = reduced_data, family = "binomial")
+```
 

### 2022-04-27 outputs/paper/paper.Rmd
introduction

--- 2022-04-24 outputs/paper/paper.Rmd
@@
 This research will focus on some social issues in the United States and how opinions of people of different religion differs.
+As the original dataset had very large amount data related to the topic of this research, decisions had to me made on which data to use.
+For the paper to be more effective on studying the influence of religion on social issues, data on popular social problems in modern society, especially in the United States have been chosen.
 

### 2022-04-28 outputs/paper/paper.Rmd
intro, abstract, result

--- 2022-04-27 outputs/paper/paper.Rmd
@@
 title: "Religion and Its Influence on Social Issues^[Code and data are available at: https://github.com/dudqo/religion_survey]"
-subtitle: "Possible factors effecting 2016 election"
+subtitle: "Agreement and Disagreement between Religion on Popular Issues in Modern Society in the U.S"
 author: 
@@
 date: "`r format(Sys.time(), '%d %B %Y')`"
-abstract: "This paper has been examined possible variables that influenced U.S citizens' choice for the 2016 U.S presidential election.
-We have found that there exists significantly different opinions regarding gender, wealth, race and age.
-This is important in a sesne that it shows thought trends as well as crucial decision indicators for voting in U.S.,With further survey at the end of this paper, we are going to explore whether this is happening only in U.S or not."
+abstract: "This paper has been examined possible influence of religion in the modern society on the popular social issues that U.S is experiencing.
+Using the U.S GSS survey data, the paper found not many significant disagreement in the popular social issues, however on the topics where disagreeemnt happen, and also on other aspects of the data, there are many important and interesting points that could be observed and is worth noting for further research."
 output:
@@
 # Introduction
-Religion has been major part of human history around the world, and for many countries, religion is something they heavily rely on even in modern days.For example, European countries such as Spain and Italy which heavily relies on catholic, and there are many middle eastern countries that heavily rely on Islam and Hinduism.
+Religion has been major part of human history around the world, and for many countries, religion is something they heavily rely on even in modern days.
+For example, European countries such as Spain and Italy which heavily relies on catholic, and there are many middle eastern countries that heavily rely on Islam and Hinduism.
 While it is reported that the number of religious people is declining quickly, religion still has big social influence in many aspects of the society in the United States, and this paper aims to research the social influence of popular religions in the United States and how the opinions of different religions differ in the context of social issues and government policies regarding them.
@@
 The dataset was obtained from U.S GSS survey, by only selecting the survey questions that are related to the topic of this research.
-The data used includes personal profile data of survey respondents such as age, gender, religion, race, income and data on questions regarding their opinion on popular social issues in the United States such as abortion, gay marriage and government policies regarding welfare, racism, and etc.
+The data used includes personal profile data of survey respondents such as age, gender, religion, race, income and data on questions regarding their opinion on popular social issues in the United States such as abortion, gay marriage and government policies regarding welfare, racism, etc.
 This research found that despite the people's different views on different religion and the differences in the religions themselves, when it comes to the popular social issues, their opinions are mostly inline with each other.
-Some different views do exists and which will be discussed in the later part of the paper.
+Some different views do exist, and which will be discussed in the later part of the paper.
 
-This report was created using `R` [@citeR] and its libraries `tidyverse` [@tidyverse], `dplyr` [@dplyr], were used to process and clean the data, and `ggplot` [@ggplot] `bookdown` [@bookdown] and `knitr` [@knitr] was used to format this report.
+This report was created using `R` [@citeR] and its libraries `tidyverse` [@tidyverse], `dplyr` [@dplyr], were used to process and clean the data, and `ggplot` [@ggplot], `bookdown` [@bookdown], `modelsummary` [@modelsummary] and `knitr` [@knitr] was used to format this report.
 
@@
 The data is from 2021 survey which was conducted online due to global pandemic by COVID19.
-For the purpose of this research, the data related to personal profiles, survey respondents' opinion on government policies and religion has been extracted from the raw survey data.
+For the purpose of this research, the data related to personal profiles, survey respondents' opinion on government policies, opinions on social issues and data on religion has been extracted from the raw survey data.
 
-The religion data has very detailed options that includes religions have very low For the simplicity of this research, protestants and Christians were grouped together under Christians and all the non-popular religion has been grouped under "other".
+The religion data has very detailed options that includes religions with very low population.
+For the simplicity of this research, protestants and Christians were grouped together under Christians and all the non-popular religion has been grouped under "other".
 
@@
 * Christian is most popular religion in the United States.
+
+Figure 1 to 6 focuses on the data of the survey respondents, such as age, race, gender, religion and income. Important points of possible limitations of this dataset to note from the figures are:
+
+* Number of survey participants of in the age group of 20-40 are very low compared to other age groups.
+* There are more female survey participants than male, however, there does not seem to be any significant impact on the result because of this.
+* As expected for a population of the United States, there are significantly more white population than any other population.
+* Also as expected, the most popular religion is Christian and Catholic, and no religion.
+
+Figure 7 to 13 are on survey respondents' opinion on many different popular social issues in modern society.
+While there is not a possible limitation shown from these plots it is worth noting that there does not seem to be big difference between the religions when it comes their opinions on social issues.
 
@@
 ```
-```{r fig4, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+```{r fig4, echo=FALSE, fig.cap = "Distribution of survey respondents' birth place"}
 reduced_data %>%
@@
 ```
-```{r fig6, echo=FALSE, fig.cap = "Distribution of religion of survey respondents by family income"}
-reduced_data %>%
-  ggplot(aes(x = relig, fill = relig)) +
-  geom_bar() +
-  labs(x = element_blank(),
-       y = element_blank(),
-       title = "Survey Respondents' Family Income by Religion",
-       ) +
-  coord_flip() +
-  theme(legend.title = element_blank())
-```
 
-```{r fig7, echo=FALSE, fig.cap = "Survey respondents' opinion on governmemt policy on improving coditions of black population"}
+```{r fig6, echo=FALSE, fig.cap = "Opinion on governmemt policy on improving coditions of black population"}
 reduced_data %>%
@@
        y = element_blank(),
-       title = "Is the Government spending enough on improving coditions of Black People?",
+       title = "Opinion on goverment spending to improve coditions of Black People",
        ) +
@@
 ```
-```{r fig8, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+```{r fig7, echo=FALSE, fig.cap = "Opinion on whether the government is spedning on welfare"}
 reduced_data %>%
@@
 
-```{r fig9, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+```{r fig8, echo=FALSE, fig.cap = "Opinion on whether the government is spending enough to assist the poor"}
 reduced_data %>%
@@
 ```
-```{r fig10, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+```{r fig9, echo=FALSE, fig.cap = "Opinion on whether the government is spending enough on social security"}
 reduced_data %>%
@@
 ```
-```{r fig11, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+```{r fig10, echo=FALSE, fig.cap = "Opinion on whether the government is spending enough to halt increasing crime rate"}
 reduced_data %>%
@@
 ```
-```{r fig12, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
-reduced_data %>%
-  drop_na(Immigration) %>%
-  ggplot(aes(x = relig, fill = Immigration)) +
-  geom_bar() +
-  labs(x = element_blank(),
-       y = element_blank(),
-       title = "Should the Government Limit Immigration to Protect U.S Culture?",
-       ) +
-  coord_flip() +
-  theme(legend.title = element_blank())
-```
 
-```{r fig13, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+```{r fig11, echo=FALSE, fig.cap = "Opinion on whether religion should have less influence"}
 reduced_data %>%
@@
 ```
-```{r fig14, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+```{r fig12, echo=FALSE, fig.cap = "Opinion on whether they support same sex marriage"}
 reduced_data %>%
@@
 ```
-```{r fig15, echo=FALSE, fig.cap = "Distribution of religion of survey respondents"}
+```{r fig13, echo=FALSE, fig.cap = "Opinion on whether they support abortion"}
 reduced_data %>%
@@
 # Results
-To study more on gender and the opinions on importance of being born as certain gender, linear model was created to study this matter.
+Looking into the personal information of the survey participants by their religion, Figure \@ref(fig:fig1) shows that distribution of age group by religion.
+It shows that number of the religious people have been declining, which suggests that the religion is becoming less relevant in the modern society.
+Figure \@ref(fig:fig2) shows the gender distribution by religion, as well as their opinion on the importance of being born as a specific gender.
+Gender data in general did not provide any useful information regarding the distribution of religion and the opinion on the importance of gender, but it can be observed that there are more female survey participants than male survey participants, but the number of non-religious people of each gender is about the same.
+Figure \@ref(fig:fig4) shows that most of the survey participants are born in the U.S and figure \@ref(fig:fig5) that income does not seem to play a role on which religion people are into.
+from figure 6 to 13, it focuses on the social issues and while there are many of them that the different religions share the same opinion on, there are some disagreements.
 
-As for the opinions on abortion, we can see an interesting result where only Christians and Catholics show significant amount of negativity towards abortion compared to other religions.
+* On the survey regarding welfare, popular answer for Christians and Catholics were that the government is spending too much, when other religions lean more towards the government is spending too little.
+* More than half of Christians, Catholics, Muslim/Islam disagree on abortion, when all other religions are very open to abortion.
+
+In general, the opinions between people of different religions does not seem to be that different, and unexpected results show that disagreements usually came from Christians and Catholics.
+However, number of survey participants from other religions were quite low and could lead to a biased result.
+
+To study more on the social issues and the influence of religion, logistics regression was used.
+The logistic regression focused on the variable with the most disagreement, which was abortion, and age, gender, race, and religion was used as predictor variables.
+
+```{r, echo=FALSE}
+abortion_mod_data <- reduced_data %>%
+  select(age_group, gender, abortion, race, relig) %>%
+  drop_na(abortion) %>%
+  mutate(across(c(age_group, gender, abortion, race, relig), as_factor))
+
+abortion_model <- glm(abortion ~ age_group + gender + race + relig, data = abortion_mod_data, family = "binomial")
+```
+
+The regression result in the Appendix shows that there is significant difference between religions coefficient estimate, which means that the there is a noticeable amount of disagreement between religions when it comes to abortion.
+However, while there are some noticeable disagreements between age groups, race and gender does not seem to have any effect on the opinions on abortion.
 
@@
 ## Summary
-In american society, general public's opinions on different religion varies significantly.
+In American society, general public's opinions on different religion varies significantly.
 While religions such as Christianity and Catholicism has been around since the beginning of history of the United States as a country, there are some religions such as Islam and Buddhism that Americans still are not very familiar with.
 With the unfamiliarity and some negative news involving a religion, the views on a specific religion be heavily influenced by stereotypes.
-We can observe from the u.S GSS survey that the opinions between people of different religion are not that different.
+We can observe from the U.S GSS survey that the opinions between people of different religion are not that different.
 
-## Different Opinions between different religions.
+## Disagreement from Christians and Catholics.
+There were some social issues that had some disagreement between religions and surprisingly, they all involved Christians and Catholics.
+While this could be due to one of the limitations of the dataset being low number of survey participants from different religions, it is quite interesting to see that the social issues popular especially in the United States had major disagreement only from the most dominant religions in the same country.
+
+## Possible decreasing influence of religion.
+Throughout the paper, there were two interesting points that can be observed from the data.
+There was noticeable disagreement between age groups regarding abortion, and the number of religious people appears to be decreasing.
+This could be an indication that the influence of religions is decreasing as number of religious people decrease.
 
 ## Weaknesses and how we should proceed in the future
-As the survey was conducted in the United States, the data lacks diversity of religions and most of survey respondents' are Christians or Catholics.
-For more accurate and in depth research, more data from people of other religions are required.
+As the survey was conducted in the United States, the data lacks diversity of religions and most of survey respondents are Christians or Catholics.
+For more accurate and in-depth research, more data from people of other religions are required.
 Also, the opinions regarding social issues could be influenced by the location they live.
-So data from different countries where Christianity or Catholicism is not the main religion could be very helpful for this research.
+So, data from different countries where Christianity or Catholicism is not the main religion could be very helpful for this research.
 
@@
 # Appendix
+```{r, echo=FALSE}
+abortion_model |>
+  modelsummary(fmt = 2, exponentiate = TRUE)
+```
 

### 2022-04-28 outputs/paper/paper.Rmd
format rmd

@@
 
-```{r fig2, fig.cap= "Distribution of survey respondents' race.", echo = FALSE}
+```{r fig2, fig.cap= "Distribution of survey respondents race.", echo = FALSE}
 reduced_data %>%
@@
 
-```{r fig3, echo = FALSE, fig.cap= "Distribution of survey respondents' gender"}
+```{r fig3, fig.cap= "Distribution of survey respondents gender.", echo = FALSE}
 reduced_data %>%
@@
 ```
-```{r fig4, echo=FALSE, fig.cap = "Distribution of survey respondents' birth place"}
+
+```{r fig4, echo=FALSE, fig.cap = "Distribution of survey respondents birth place"}
 reduced_data %>%
@@
 ```
+
 ```{r fig7, echo=FALSE, fig.cap = "Opinion on whether the government is spedning on welfare"}
@@
 ```
+
 ```{r fig9, echo=FALSE, fig.cap = "Opinion on whether the government is spending enough on social security"}
@@
 ```
+
 ```{r fig10, echo=FALSE, fig.cap = "Opinion on whether the government is spending enough to halt increasing crime rate"}
@@
 ```
+
 ```{r fig12, echo=FALSE, fig.cap = "Opinion on whether they support same sex marriage"}
@@
 ```
+
 ```{r fig13, echo=FALSE, fig.cap = "Opinion on whether they support abortion"}
